### 2024-11-27 other/datasheet/datasheet_template.qmd
add datasheet and appendix

---
title: "Datasheet for 'A dataset'"
author: 
  - Colin Sihan Yang
thanks: "Code and data are available at: [https://github.com/hoodiexxx/lol_win_rate_prediction](https://github.com/hoodiexxx/lol_win_rate_prediction)."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
format: pdf
number-sections: true
bibliography: ../../paper/references.bib
---

# Datasheet for BLG Player Performance Dataset

## Motivation

1. *For what purpose was the dataset created?*  
   The dataset was created to predict player performance for Bilibili Gaming (BLG) in the League of Legends Pro League (LPL). It aims to assist teams in making data-driven decisions on player strategies, performance optimization, and resource allocation.

2. *Who created the dataset?*  
   The dataset was created by Riot Game and oracleselixir @oracleselixir_oracles_nodate.

## Composition

1. *What do the instances represent?*  
   Each instance represents an individual game played by a BLG player, including in-game metrics and results.

2. *How many instances are there in total?*  
   The dataset contains approximately 470 instances, corresponding to all games played by BLG in the 2024 LPL season.

3. *Does the dataset contain all possible instances?*  
   The dataset is a complete record of BLG games for the selected season. However, it does not include games from other teams or non-professional matches.

4. *What data does each instance consist of?*  
   Each instance includes metrics such as vision score, total gold, total CS, damage to champions, KDA, game duration, and match outcomes.

5. *Is there a label or target associated with each instance?*  
   Yes, the target variable is the Kill-Death-Assist ratio (KDA).

6. *Is any information missing from individual instances?*  
   No significant data is missing. All in-game metrics are complete.

7. *Are relationships between instances made explicit?*  
   No explicit relationships between instances are defined. Each instance is independent.

8. *Are there recommended data splits?*  
   Yes, the dataset is split into training (80%) and testing (20%) sets.

9. *Are there errors, noise, or redundancies?*  
   Minor inconsistencies may exist due to variations in game logs, but these have been minimized during preprocessing.

10. *Is the dataset self-contained?*  
    Yes, the dataset is self-contained and does not rely on external resources.

11. *Does the dataset contain confidential data?*  
    No, all data is derived from publicly available match records.

12. *Does the dataset contain offensive or sensitive data?*  
    No, the dataset is free from offensive or sensitive content.

13. *Does the dataset identify subpopulations?*  
    Yes, the dataset categorizes players by role (e.g., Top Lane, Jungle).

14. *Is it possible to identify individuals?*  
    No, the dataset only includes player names and their roles without sensitive personal information.

15. *Does the dataset contain sensitive data?*  
    No, the dataset does not include sensitive attributes like race, religion, or financial data.

16. *Any other comments?*  
    None.

## Collection Process

1. *How was the data acquired?*  
   The data was extracted from publicly available match records and APIs such as Riot Games' data API.

2. *What mechanisms were used to collect the data?*  
   Data was collected using API scripts and verified for completeness and accuracy.

3. *If the dataset is a sample, what was the sampling strategy?*  
   The dataset is not a sample; it is a complete record of BLG's 2024 matches.

4. *Over what timeframe was the data collected?*  
   Data was collected during the 2024 LPL season.

5. *Were any ethical review processes conducted?*  
   Ethical approval was not required as the data is publicly available.

6. *Did you collect the data directly?*  
   Yes, data was collected from Riot Games' API and match records.

7. *Were individuals notified about the data collection?*  
   Not applicable, as the data pertains to publicly available information.

8. *Did individuals consent to data collection?*  
   Not applicable, as the data pertains to publicly available information.

9. *Any other comments?*  
    None.

## Preprocessing, Cleaning, and Labeling

1. *Was any preprocessing done?*  
   Yes, data was cleaned to remove missing values and calculate derived metrics like KDA.

2. *Was raw data saved?*  
   Yes, the raw data is archived for future use.

3. *Is the preprocessing software available?*  
   Yes, preprocessing scripts are available in the accompanying GitHub repository.

4. *Any other comments?*  
   None.

## Uses

1. *Has the dataset been used for any tasks already?*  
   Yes, it was used to build a Bayesian mixed-effects model for predicting player KDA.

2. *What other tasks could the dataset be used for?*  
   The dataset could be used for performance analysis, team strategy optimization, and meta-analysis of in-game metrics.

3. *Are there tasks for which the dataset should not be used?*  
   The dataset should not be used for any non-esports purposes, such as identifying personal information.

4. *Any other comments?*  
   None.

## Maintenance

1. *Who will maintain the dataset?*  
   No one

2. *How can the dataset owner be contacted?*  
   N/A

3. *Will the dataset be updated?*  
    N/A

4. *Any other comments?*  
   None.

\newpage

# References